draft 1

### 2015-08-02 final write up/draft.rmd
wkend work

@@

## 1.1 Research Motivations

* thousands of consulting companies globally
* range from single consultant to 10's of 1000's of employees
* employees broad range of people from fresh graduates, to founders with decades of experience
* internal environment: worker feelings of morale and security can be heavily influenced by success and profitability of their projects. Not to mention possibilities for pay rises and promotions which all come together to create a positive thriving culture. 
* harsher economic climate, particularly important to get competitive edge
* find a source explaining how companies create their fixed fees - intuition, experience
* find a source saying how badly it goes!
* in the case study (and is typical of smaller organisations??), the CRM is an untapped source of information storing
    * employee hours
    * client information/characteristics
    * invoicing
    * employee costs
    * employee charge out rates
    * project information
* currently the case study company uses a widely popular CRM, where they can output information but cannot perform modelling or analysis:
    * look up one job at a time
    * basic summaries of overall hours spent for the month

>>>>

****

@@

## 1.1 Research Motivations - 4.5 pages

description of problem:

Project decision makers have struggled with forecasting project costs accurately for decades, across many industries. A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% (Flyvbjerg, 2006). A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average (Flyvbjerg and Budzier, 2011). Complex longer term projects are necessary to produce  railways, houses, websites, and movies which involve professionals across a many industries. Depending on the industry, consulting companies are enlisted to complete portions of a project, and some of the financial risk of the overall project may be transferred to the consultant. This is achieved by requiring a fixed flat fee from the consultant, meaning they have committed to perform their role for the nominated price. In a similar way to infrastructure and IT projects, consultants are susceptible to underestimating the amount of their time that will be spent in their role, and must wear the cost of extra employee hours spent. For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.

The ramifications of unprofitable projects in a private business are multifold. Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery. Employees also experience stress attempting to complete the project within a disappearing budget and may perform lower quality work in exchange for speed. This is unfortunate if the reality is that the project fee was under estimated. Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would improve a business' standing. The financial risk taken on fixed flat fee projects (and disappointing results) may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area. Clearly, failure to predict fixed fee project costs can negatively affect a business' growth opportunities and staff morale.

Why its not easily solvable:

The challenge for consultants to reduce the risk of a project going over budget, in terms of hours spent, is not easily solved. If a job does go over budget, the business experiences reduced profits or  even losses as they are contractually bound to complete the work. The reason this task is challenging is that complex projects are always different from each other, and similar work may take significantly different amounts of time. For example, an ad filmed in a remote location for chocolate will cost significantly more than an ad filmed on an exisitng set, even though the product structure is almost identical. A fixed product price, which you could apply to repetive sales such as appliances, is inappropriate. One option is charging for consultants' time by the hour; a zero risk fee structure. However in many industries this is not competitive enough. Other consultancies may be willing to offer a fixed flat fee for the same work, reducing financial risk for the client, and luring the client away. Another fee strategy may be to track worker hours, and stop work once the consulting fee has been depleted. Then, negotiate further fees, or variations, with the client and persuade them to accept. This also satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective. However, if the project has been following a predictable process, consultants often prefer to wear losses or low profits to preserve the relationship with their client. Consultant's may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry over creating financial friction. Clearly, there is no straightforward solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects. In some indusries, fixed flat fee structures are expected and business' must conform to survive.

Solutions:

Previous research has been done on the prediction of final costs of a complex project with varying success. In the prediction of IT software development, formal estimation methods have existed for many years, however expert judgment is still the dominant method and available evidence does not suggest the models improve on expert judgment (Jorgensen and Shepperd, 2007). Construction industry research has used models ranging from neural networks to case based reasoning to regression to predict final building costs. However, often expert judgment was not mobilised, and the models were not comprehensive enough to be used in industry. Manager's experience and intuition is still the dominant method for construction cost calculation.

- research has gone into targeting common repeated points in a projects timeline where things can go awry. reference
- example: find research, IT projects? nothing in engineering or film?
- Although no solutions are available, the use of reference class forecasting is being pioneered in england for large government infrastructure
- method for fee pricing has remained the same for decades - detailed analysis of project, add up time based on tasks, allow buffer time for error.

Reference class forecasting
* predicting cost of construction from other perspectives - builders.

Other industries
* predicition of building construction costs - statistical analysis

Potential solution:
* stat analysis like reference class forecasting and building construction except analysis performed internally using the company's historical data on past projects.

case study:

This project undertakes a case study on a consulting company in the construction industry. Specifically, the

* in the case study (and is typical of smaller organisations??), the CRM is an untapped source of information storing
    * employee hours
    * client information/characteristics
    * invoicing
    * employee costs
    * employee charge out rates
    * project information
* currently the case study company uses a widely popular CRM, where they can output information but cannot perform modelling or analysis:
    * look up one job at a time
    * basic summaries of overall hours spent for the month
